---
title: "Web scraping in R"
author: "H33M"
date: "`r Sys.Date()`"
output: html_document
---
 What is Web Scraping?
Web scraping is a technique for converting the data present in unstructured format (HTML tags) over the web to the structured format which can easily be accessed and used.
*rvest library* is famous for web scraping and harvesting data from a web page. 
## Scraping a web page in R
install and load the rvest library package.
```{r load the package}
#install using install.packages("rvest")
library(rvest)# load the package
```
in order to start parsing that data, first request it from the computer server that contains it.
in rvest package function that serves this purpose is the *read_html function*. read_html function takes a web URL as an argument.
```{r read the html}
simple <- read_html("https://dataquestio.github.io/web-scraping-pages/simple.html")
simple
```
*read_html function* returns a list containing the html and css structures.
in order to capture the _text_ use the *html_text function* while the _html tags_ use the *html_nodes function*. the html tags are shown in our text using the <p> tags.
```{r capture the text and tags}
simple %>% 
  html_nodes("p") %>% 
  html_text()
```
Once we have all of these nodes, we can pass the output of html_nodes() into the html_text() function. We needed to get the actual text of the <p> tag, so this function helps out with that.

These functions together form the bulk of many common web scraping tasks. In general, web scraping in R (or in any other language) boils down to the following three steps:

1. Get the HTML for the web page that you want to scrape
2. Decide what part of the page you want to read and find out what HTML/CSS you need to select it
3. Select the HTML and analyze it in the way you need

Web scraping on a national weather service website as shown below.
```{r web scraping weather service website}
forecasts<- read_html("https://forecast.weather.gov/MapClick.php?lat=37.7771&lon=-122.4196#.Xl0j6BNKhTY") %>% # load the web URL for scraping.
  html_nodes(".temp") %>% # extract the html tags using html_nodes
  html_text() #capture the text using html_text
forecasts
```
Get the data into the format that we want or format that can be used. for example:
```{r}
library(readr)
parse_number(forecasts) # extract the temparatures 
```

Scraping IMDb website for  100 most popular feature films released in 2016. Specify the web url using url variable as shown below.
```{r IMDb web scraping}
#specify the web url
web_url <- "http://www.imdb.com/search/title?count=100&release_date=2016,2016&title_type=feature"
#Reading the html from the website
web_page <- read_html(web_url)
```
We'll scrap the following;
rank, title, description, runtime, genre, rating, metascores, votes, gross_earned_in_mil, director, actor.
Start by scraping the Rank field. Once you know the css selector that contains the rankings, use the codes below to extract the information for all rankings for these movies.
```{r web scrap the rank field}
rank_data <- web_page %>% 
  html_nodes(".text-primary") %>% # extract html tags
  html_text()
rank_data %>% 
  head()
```
convert the data into the format that is desirable. for this data, preprocess it into numeric data. 
```{r convert into numeric data}
rank_data <- as.numeric(rank_data)
head(rank_data)
```
do the same for title. Start by checking the css selector for title.
```{r web scraping for title section}
title_data <- web_page %>%  html_nodes(".lister-item-header a") %>% 
  html_text()
title_data %>% 
  head()
```
web scraping and data preprocessing for description.
```{r web scraping for description}
# web scraping
desc_data <-web_page %>%  
  html_nodes(".ratings-bar+ .text-muted") %>% 
  html_text()
#data preprocessing
desc_data <- gsub("\n"," ", desc_data)
head(desc_data)
```

web scraping for Runtime, genre,  Rating, Metascore, Votes, Gross_Earning_in_Mil , Director and Actor data.
```{r web scraping runtime section}
## web scraping the runtime section

runtime_data <-web_page %>%
  html_nodes('.text-muted .runtime') %>% 
  html_text()
#preprocess the data
runtime_data <- gsub("min", " ", runtime_data)
runtime_data <- as.numeric(runtime_data)
head(runtime_data)
```

```{r web scraping genre section}
## web scraping the Genre section
genre_data <- web_page %>% 
  html_nodes(".genre") %>% 
  html_text()
# preprocess the data
# remove \n
genre_data<- gsub("\n", " ", genre_data)
# remove extra spaces
genre_data <- gsub(" ", "", genre_data)
# taking the first genre only
genre_data<- gsub(",.*", " ", genre_data)
#convert to factors
genre_data<- as.factor(genre_data)
head(genre_data)

```
```{r web scraping the rating section}
#web scraping the rating section
rating_data <- web_page %>% 
  html_nodes(".ratings-imdb-rating strong") %>% 
  html_text()
#preprocessing the data for use
rating_data<- as.numeric(rating_data)
head(rating_data)
```
```{r scraping the votes section}
votes_data<- web_page %>% 
  html_nodes(".sort-num_votes-visible span:nth-child(2)") %>% 
  html_text()
# preprocess the data
#remove commas
votes_data <- gsub(",", "", votes_data)
#convert to numeric
votes_data<- as.numeric(votes_data)
head(votes_data)
```
```{r web scraping the director section}
directors_data <- web_page %>%
  html_nodes('.text-muted+ p a:nth-child(1)') %>%
  html_text()
# data preprocessing: converting into factor
directors_data <- as.factor(directors_data)
head(directors_data)

```

```{r web scraping the actor section}
actor_data <- web_page %>% 
  html_nodes('.lister-item-content .ghost+ a') %>%
  html_text()
#convert into factors
actor_data <- as.factor(actor_data)
head(actor_data)
```
```{r web scraping for metascore section}
metascore_data <- web_page %>% 
  html_nodes(".metascore") %>% 
  html_text()
#preprocess the data
# remove extra space in the data
metascore_data <- gsub(" ", "", metascore_data)
length(metascore_data) # some data are missing for metascore_data therefore will need to be replaced with NAs.the positions without metascores are 63,66 and 24
for(i in c(24,63,66)){
  a<- metascore_data[1:(i-1)]
  b <- metascore_data[i:length(metascore_data)]
  metascore_data <- append(a, list("NA"))
  metascore_data<- append(metascore_data, b)
}
#convert to numeric data
metascore_data<- as.numeric(metascore_data)
head(metascore_data)
summary(metascore_data)
```
```{r gross earnings variable web scraping}
gross_earning_data <- web_page %>% 
  html_nodes('.ghost~ .text-muted+ span') %>% 
  html_text()
# data preprocessing
# remove M
gross_earning_data<- gsub("M","", gross_earning_data)
#remove $ sign
gross_earning_data <- substring(gross_earning_data,2,6 )

length(gross_earning_data) # this variable have some missing data thus need to be replaced with NAs

for(i in c(24, 37,51,58, 60, 63,65,66)){
  a<- gross_earning_data[1:(i-1)]
  b<- gross_earning_data[i:length(gross_earning_data)]
  gross_earning_data<- append(a, list("NA"))
  gross_earning_data<- append(gross_earning_data,b)
}
# convert into numeric
gross_earning_data<- as.numeric(gross_earning_data)

```
Combine the eleven variables to form a data frame. 
```{r combining the data to form a dataframe}
movies_df <- data.frame(
  Ranks= rank_data,
  Title= title_data,
  Description= desc_data,
  Runtime= runtime_data,
  Genre= genre_data,
  Rating= rating_data,
  Metascore= metascore_data,
  Gross_Earning_in_mils= gross_earning_data,
  Votes= votes_data,
  Director= directors_data,
  Actor= actor_data
)
View(movies_df)
```
Analysing the scraped data can also be done as follows.
```{r Analysing the scraped data}
library(ggplot2)
#plot the Runtime filled with Genre
qplot(data = movies_df, Runtime, fill= Genre, bins = 30)
#which movie from which genre have the longest runtime
ggplot(data = movies_df, aes(x= Runtime, y= Rating))+ 
  geom_point(aes(size= Votes, color= Genre))
#in the Runtime of 130-160 mins, which genre has the highest votes?
ggplot(data= movies_df, aes(x= Runtime, y= Gross_Earning_in_mils))+
  geom_point(aes(size = Rating, color= Genre))
```


